---
title: "admixture_results"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(genio)
#devtools::install_github('StoreyLab/popkin')
library(popkin)
library(ochoalabtools)
```

Load admixture proportions `Q` and aligned IDs from `fam`.
Then load and align demographics `data` to `fam`; `fam` is not used after this.
```{r}
# assume we're here:
# /datacommons/ochoalab/ssns_gwas/nephrotic.syndrome.gwas_proprocessing_202205/admixture/

Q <- read_matrix("ssns_gwas_maf.5", ext = 'Q') 
fam <- read_fam( '../ssns_gwas_maf' )

data <- read_csv( "../pheno_excelmerge.csv", show_col_types = FALSE )
data <- data %>% filter(AcquisitionNumber %in% fam$id) 
data <- data[match(fam$id, data$AcquisitionNumber),]
```

Merge minor groups, then order individuals by race:
```{r}
# merge these into Other, to simplify plot only
data$RACE[ data$RACE == 'Mixed' ] <- 'Other'
data$RACE[ data$RACE == 'Unknown' ] <- 'Other'

race_order <- c("Black", "White", "Asian", "Hispanic", "Other")
indexes <- order( match( data$RACE, race_order ) )
# order both Q and data so they stay aligned
Q <- Q[ indexes, ]
data <- data[ indexes, ]
```

Automatically label ancestries based on race in which they're most common:
```{r}
# desired ancestry labels and correspondence to race
# (Asian1/2 had to see prelim plots to get association)
ancestry_order <- c('African', 'European', 'South Asian', 'East Asian', 'Native American')
ancestry_to_race <- c('Black', 'White', 'Asian1', 'Asian2', 'Hispanic')

# get best race assignment
anc_names <- admix_label_cols( Q, labs = data$RACE )
# manually replace race with corresponding ancestry labels
anc_names[ match( ancestry_to_race, anc_names ) ] <- ancestry_order
# store in Q matrix now
colnames( Q ) <- anc_names
# reorder further to get ancestries in desired order
indexes <- match( ancestry_order, anc_names )
Q <- Q[ , indexes ]
```

Do some manual reordering within race groups for some main ancestry
```{r}
reorder_race_anc <- function( Q, data, race, ancestry ) {
    indexes <- which( data$RACE == race )
    indexes2 <- order( Q[ indexes, ancestry ], decreasing = TRUE )
    Q[ indexes, ] <- Q[ indexes[ indexes2 ], ]
    data[ indexes, ] <- data[ indexes[ indexes2 ], ]
    return( list( Q = Q, data = data ) )
}
# apply to each race/ancestry set
obj <- reorder_race_anc( Q, data, 'Black', 'African' )
obj <- reorder_race_anc( obj$Q, obj$data, 'White', 'European' )
obj <- reorder_race_anc( obj$Q, obj$data, 'Asian', 'South Asian' )
obj <- reorder_race_anc( obj$Q, obj$data, 'Hispanic', 'Native American' )
obj <- reorder_race_anc( obj$Q, obj$data, 'Other', 'European' )
Q <- obj$Q
data <- obj$data
```

Further separate less from more admixed individuals in each of the 3 major ancestry.
This is the key filtering step that enables the ancestry subanalyses!
```{r}
# threshold for ancestry subanalyses
q_cut <- 0.8
data$RACE[ data$RACE == 'Black' & Q[ , 'African' ] <= q_cut ] <- 'Black-Adm.'
data$RACE[ data$RACE == 'White' & Q[ , 'European' ] <= q_cut ] <- 'White-Adm.'
data$RACE[ data$RACE == 'Asian' & Q[ , 'South Asian' ] <= q_cut ] <- 'Asian-Adm.'

# for reference, race/ethnicity clusters by subclassification
table( data$RACE, data$DIAGNOSIS )
##            Control NS UNCLASSIFIED SRNS SSNS
## Asian          248               0   14  329
## Asian-Adm.      67               0    2   18
## Black          301               5   41  148
## Black-Adm.     148               0   29   23
## Hispanic         2               6   22   29
## Other            5               2    4   19
## White          274               1   75  136
## White-Adm.       4               0    6   23
table( data$RACE )
##      Asian Asian-Adm.      Black Black-Adm.   Hispanic      Other      White 
##        591         87        495        200         59         30        486 
## White-Adm. 
##         33 
table( data$DIAGNOSIS )
## Control NS UNCLASSIFIED            SRNS            SSNS 
##    1049              14             193             725 
nrow( data ) # [1] 1981
```

Make structure plot!
```{r fig.width=20, fig.height=4}
fig_start( 'admix-k5', mar_t = 0.5, mar_b = 4, height = 4, width = 9 )
plot_admix(
    Q,
    labs = data$RACE,
    labs_cex = 0.7,
    labs_even = TRUE,
    labs_line = 2,
    xlab_line = 3,
    leg_cex = 0.5,
    leg_width = 0.1
)
fig_end()
```

### PCA analysis

Load PCs, align to Q/data.
```{r}
obj <- read_eigenvec("../pca/PCA10")
pcs <- obj$eigenvec
fam <- obj$fam

indexes <- match( data$AcquisitionNumber, fam$id )
fam <- fam[ indexes, ]
pcs <- pcs[ indexes, ]
data$PC1 <- pcs[, 1]
data$PC2 <- pcs[, 2]
data$PC3 <- pcs[, 3]

# plot looks better if race is a factor in desired order
data <- data %>% mutate( RACE = fct_inorder( RACE ) )
```

Make PCA plot!
```{r}
fig_start( 'admix-k5-pca', height = 6, width = 7 )
ggplot(data, aes(x = PC1, y = PC2, color = RACE)) + geom_point() + theme_classic() + labs(color = 'Race') +
    scale_color_brewer(palette = "Spectral") # Paired
fig_end()
```

### write outliers into files and calculate allele frequency

```{r}
write_ids <- function( data, race, name ) {
    # get data to output
    x <- data$AcquisitionNumber[ data$RACE == race ]
    x <- paste0( x, " ", x )
    # set filepath
    file <- paste0( "../allele_freq/ssns_maf_", name, ".txt" )
    # write it out!
    write_lines( x, file )
}

write_ids( data, 'Asian', 'sa' )
write_ids( data, 'Black', 'black' )
write_ids( data, 'White', 'white' )
```

